The Practical Application of MST

### 2021-02-14 article.Rmd
Draft sections on data sources
Section on mstR

I'll assume some familiarity with IRT or Item Response Theory (at least with Rasch analysis), although I provide a brief summary below. Also some competence with R programming would be useful to understand the examples given. For background information and primers on these subjects I provide references.

My main focus is the development of tests for language learners, specifically English language learners. The theory and data processing involved can easily be transferred to other domains that require measurement of latent ability. The IRT model that I focus on is the 1PL (or Rasch) model. For readers interested in extending analysis to more complex models, see [@magis2017computerized].

## Introduction to Item Response Theory
The underlying theoretical basis for MST is the branch of psychometrics known as Item Response Theory, or IRT. The theory proposes that there are *latent traits* (for example language ability) that can be measured using a set of tests delivered to the test candidate. These test items are typically scored as correct (1) or incorrect (0) - and it these dichotomous test types that we focus on here. The response to the test items can then be use to estimate the latent trait by fitting an IRT model.

### The Rasch Model
The simplest such model, the so-called *Rasch model*, relies on a single variable parameter, the difficulty parameter (usually designated as $b$). For this reason it is also known as the one-parameter logistic model (1PL). The *item response function* (IRF) for this model uses a logistic function to model the probability of a correct response for candidate $i$ taking test item $j$:

$$
Pr(X_{ij} = 1|\theta_{i},b_{j}) = \frac{exp(\theta_{i} - b_{j})}{1+exp(\theta_{i} - b_{j})}
$$
Here $X$ is the item response (either 1 or 0), $\theta$ represents the latent ability being measured, and $b$ represents the difficulty of the item. Certain assumptions are made by the model, most importantly:

@@

![Logistic Function](images/icc.png)
There is an equivalence between $\theta$ and $b$ such that a candidate with mean ability ($\theta = 0$) would be expected to get item US065_1 (green) correct with a probability of 0.28. The less difficult item US070_1 (black) would see a probability of around 0.95 for a correct response. Item US070_1 (red) is somewhere in between the two in terms of difficulty. You could also read this as stating that candidates who have a $\theta$ of 0.82 show a 50-50 chance of getting US065_1 correct. The difficulty parameter $b$ is set at this point, so we'd assign a difficulty of 0.82 to item US065_1 in our test. And assuming a normal distribution of candidates, items with difficulty 0 would expect to be answered correctly by 50% of candidates.

## IRT for Mulitstage Test Design
IRT is used to inform calibration of the test items after they've been developed by subject matter experts, as well as construction of the modules and routing between them. Scoring of the test can be derived from the $\theta$ estimates. Typically the initial module (known as the routing module) will have an overall difficulty of 0 (ie. a mix of easy and hard items), with subsequent routing directing candidates to easier or more difficult modules in order to 'hone in' on an accurate $\theta$ estimate at the end of the test.

@@

## MST Design Considerations?
	- Terminology
	- Benefits
	- Design

@@

	- Design test and validation

6. Trialling and Test Validation
	- Item Writing and Intial Validation
	- Trialling
	- Import and validation of trial data
	- Updating the item bank
	- Other outputs

>>>>

I'll assume some familiarity with Item Response Theory (IRT) - and the Rasch model. Also some competence with R programming would be useful to understand the examples given. For background information and primers on these subjects I provide references.

My main focus is the development of tests for language learners, specifically English language learners. The theory and data processing involved can  be transferred to other domains that require measurement of latent ability. The IRT model that I focus on is the 1PL (or Rasch) model. For readers interested in extending analysis to more complex models, see [@magis2017computerized].

## Introduction to Item Response Theory

The underlying theoretical basis for MST is the branch of psychometrics known as Item Response Theory, or IRT. The theory proposes that there are *latent traits* (for example language ability) that can be measured using a set of tests delivered to the test candidate. These test items are typically scored as correct (1) or incorrect (0) - and it these dichotomous test types that we focus on here. The response to the test items can then be use to estimate the latent trait by fitting an IRT model.

### The Rasch Model

The simplest such model, the so-called *Rasch model*, relies on a single variable parameter, the difficulty parameter (usually designated as $b$). For this reason it is also known as the one-parameter logistic model (1PL). The *item response function* (IRF) for this model uses a logistic function to model the probability of a correct response for candidate $i$ taking test item $j$:

$$
Pr(X_{ij} = 1|\theta_{i},b_{j}) = \frac{exp(\theta_{i} - b_{j})}{1+exp(\theta_{i} - b_{j})}
$$

Here $X$ is the item response (either 1 or 0), $\theta$ represents the latent ability being measured, and $b$ represents the difficulty of the item. Certain assumptions are made by the model, most importantly:

@@

![Logistic Function](images/icc.png)

There is an equivalence between $\theta$ and $b$ such that a candidate with mean ability ($\theta = 0$) would be expected to get item US065_1 (green) correct with a probability of 0.28. The less difficult item US070_1 (black) would see a probability of around 0.95 for a correct response. Item US070_1 (red) is somewhere in between the two in terms of difficulty. You could also read this as stating that candidates who have a $\theta$ of 0.82 show a 50-50 chance of getting US065_1 correct. The difficulty parameter $b$ is set at this point, so we'd assign a difficulty of 0.82 to item US065_1 in our test. And assuming a normal distribution of candidates, items with difficulty 0 would expect to be answered correctly by 50% of candidates.

## IRT for Mulitstage Test Design

IRT is used to inform calibration of the test items after they've been developed by subject matter experts, as well as construction of the modules and routing between them. Scoring of the test can be derived from the $\theta$ estimates. Typically the initial module (known as the routing module) will have an overall difficulty of 0 (ie. a mix of easy and hard items), with subsequent routing directing candidates to easier or more difficult modules in order to 'hone in' on an accurate $\theta$ estimate at the end of the test.

@@

## MST Design Considerations

I mentioned earlier that multi-stage test designs comprise a number of modules arranged into stages. The question then becomes, how many stages do I need, and how many modules in each stage? The simplest design is a 1-2-3 design, laid out like this:

<<TODO _ 1-2-3 design >>

However, this design may not be sufficient for your needs. You need to consider whether there are groups of items that naturally fit into a particular stage (for example, seperating reading or listening tasks into seperate stages). You also need to consider how many items per stage you want to have. Also, in order for the routing between modules to produce acceptable results, you need to consider the relative difficulty of each module.

Normally the so-called routing module in stage 1 (Module A) would contain a mix of items with a mean difficulty of 0 (in other words 'medium' difficulty). Candidates that are placed in the lower ability range at the end of this module will be routed to the 'easy' module in stage 2 (Module B), and those in the upper ability range to the 'hard' module (Module C).

@@

An initial simulation of your test to try out designs can be done by using the `mstR` library (see @magis2017computerized). Here the aim is to select a set of designs and see if there is an optimal design. It will also help understand how the scoring of the test should be implemented, including the number correct (NC) scores that will be used for routing, and the ability estimates for final scoring and reporting.

The `mstR` library can take an existing set of item responses, or you can use it to generate a set of randomised simulated item responses. This latter ability is useful in the pre-trial stage where we're interested in finding out how routing between modules should be established, or to validate assumptions in the test design.

1. Trialling and Test Validation
	- Item Writing and Intial Validation
	- Trialling
	- Import and validation of trial data
	- Updating the item bank
	- Other outputs
